Computing Entropies With Nested Sampling

### 2017-08-14 paper/response.tex

\documentclass[a4paper, 12pt]{article}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{dsfont}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage[left=2cm, right=2cm, bottom=3cm, top=2cm]{geometry}
\usepackage{natbib}
\usepackage{microtype}
\usepackage{amssymb}
\usepackage{color}

\definecolor{orange}{rgb}{0.7, 0.25, 0}
\definecolor{green}{rgb}{0, 0.5, 0}
\definecolor{darkblue}{rgb}{0, 0, 0.7}

\renewcommand{\quote}{\em \color{orange}}

@@

{\quote
Estimation of entropy (and other information-theoretic quantities) and estimation of the uncertainties of the estimates is a notoriously difficult problem.  I personally, would have benefited greatly from having access, or knowledge of, such a methodology years ago.  The paper should definitely be published after some minor edits.}

That's why I started thinking about this problem...

{\quote
1. The author presents a methodology for computing entropies, and other information-theoretic quantities, using Skilling's nested sampling algorithm.  It should be noted that Skilling himself advocates the estimation of the "entropy", I believe that it is the KL-divergence, as the estimate of the uncertainty in the estimated evidence.  This should be mentioned, and possibly discussed within the context of this present paper.}

I have mentioned that NS computes the prior-to-posterior KL divergence, and
that this has both an interesting interpretation and is used as a termination
criterion. This was added to the first paragraph of Section 3.

@@

{\quote
9. It wasn't clear to me how the distance function (25) was chosen and why.}

It is just the Euclidean distance between two data vectors, each of which
contains 100 co-ordinates. I added a sentence to the paper to clarify this.

{\quote
10. I had some questions about how the algorithm was conceived.  I had a difficult time envisioning how the blue distribution played the role of the prior, and why.  Perhaps some explanation is necessary here since NS is usually presented to as to assume that parameter changes were implemented to maintain a uniform prior.}

I know a few groups use that convention (transform the parameter space so
the prior becomes uniform), but I never have, and there's no reason to do
that in general. As long as it's possible to explore $\pi$, then everything
will be fine.

@@

{\quote
11.  It was not entirely clear to me how the algorithm was used to estimate the entropy of a Gaussian.  How was the posterior (24) arrived at.  The prior is normal.  But what about the likelihood?  And then how was the distance measure (25) computed?}

I have added an extra paragraph with the intent of explaining this in more
detail.

{\quote
There are a number of typos:

1. End of Section 1: "I sometimes write general formulae are written ..."}

Fixed.

{\quote
2. Missing citation in first sentence of Section 6}

I removed that citation.

{\quote
3. After (47): question mark missing in case ii)}

Fixed.

### 2017-08-14 paper/ms.tex
Finished Kevin's comments

It is important to distinguish between $H(\x | \d)$ and the
entropy of $\x$ given a particular value of $\d$, which might be written
$H(\x | \d=\d_{\rm observed})$.
The former measures the degree to which one question answers
another {\em ex ante}, and is a function of two questions.
The latter measures the degree to which {\revision a question remains unresolved
after conditioning on a specific statement ({\em ex post})},
and is a function of a question and a statement.

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=0.5\linewidth]{figures/algorithm.pdf}
\caption{To evaluate the log-probability (or density) of the blue
probability distribution {\revision near} the red point, Nested Sampling can be used,
with the blue distribution playing the role of the ``prior'' in NS, and the
Euclidian distance from the red point (illustrated with red contours)
playing the role of the negative log-likelihood. Averaging over selections
of the red point gives an estimate of the entropy of the blue distribution.
\label{fig:algorithm}}
\end{figure}

@@

The joint entropy $H(\ln \ytot, \ln \ztot)$ was also estimated by using
a Euclidean distance function and a tolerance of $10^{-3}$. The result was
\begin{align}
\left<\depth_{(\ln \ytot, \ln \ztot)}\right> &= 16.03 \pm 0.11.
\end{align}
Applying the log-volume correction yields a differential entropy estimate of
$H(\ln \ytot, \ln \ztot) = 3.40 \pm 0.11$ nats.
Therefore, the mutual information is
\begin{align}
I(\ytot; \ztot) &= H(\ln\ytot) + H(\ln\ztot) - H(\ln\ytot, \ln\ztot)\\
                &= 1.130 \pm 0.082 \textnormal{ nats}.
\end{align}
The variance of this mutual information estimate was reduced by
using a common sequence of reference points for the marginal and joint entropy
runs, an instance of the `common random numbers'
technique\footnote{\tt http://demonstrations.wolfram.com/TheMethodOfCommonRandomNumbersAnExample/}.

>>>>

It is important to distinguish between $H(\x | \d)$ and the
entropy of $\x$ given a particular value of $\d$, which might be written
$H(\x | \d=\d_{\rm obs})$ (`obs' for observed).
The former measures the degree to which one question answers
another {\em ex ante}, and is a function of two questions.
The latter measures the degree to which {\revision a question remains unresolved
after conditioning on a specific statement ({\em ex post})},
and is a function of a question and a statement.

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=0.5\linewidth]{figures/algorithm.pdf}
\caption{To evaluate the log-probability (or density) of the blue
probability distribution {\revision near} the red point, Nested Sampling can be used,
with the blue distribution playing the role of the ``prior'' in NS, and the
Euclidean distance from the red point (illustrated with red contours)
playing the role of the negative log-likelihood. Averaging over selections
of the red point gives an estimate of the entropy of the blue distribution.
\label{fig:algorithm}}
\end{figure}

@@

The joint entropy $H(\ln \ytot, \ln \ztot)$ was also estimated by using
a Euclidean distance function and a tolerance of $10^{-3}$. The result was
\begin{align}
\left<\depth_{(\ln \ytot, \ln \ztot)}\right> &= 16.03 \pm 0.11.
\end{align}
Applying the log-volume correction yields a differential entropy estimate of
$H(\ln \ytot, \ln \ztot) = 3.40 \pm 0.11$ nats.
Therefore, the mutual information is
\begin{align}
I(\ytot; \ztot) &= H(\ln\ytot) + H(\ln\ztot) - H(\ln\ytot, \ln\ztot)\\
                &= 1.130 \pm 0.082 \textnormal{ nats}.
\end{align}
The variance of this mutual information estimate was reduced by
using a common sequence of reference points for the marginal and joint entropy
runs. This is an instance of the `common random numbers'
technique\footnote{\tt http://demonstrations.wolfram.com/TheMethodOfCommonRandomNumbersAnExample/}.
{\revision In short, for those reference particles with a large depth
in terms of the $\ln \ytot$ distance function, the depth in terms of the
$(\ln \ytot, \ln \ztot)$ distance function was also large. Since the
mutual information involves a difference of the averages of these depths,
it is more accurate to compute the average difference 
(i.e. compute the two depths using the same reference particle, to compute
the difference between the two depths) than the difference of the averages.
}
